On the Complexity of Best Arm Identification

### 2017-05-28 Complexity of Best Arm Identification/kl_complexity.Rmd
Add conclusion to KL Complexity chapter

\begin{equation*}
\forall x \in [0,1], \quad \der(x,1-x) = \der(1-x,x) \geq \log \frac{1}{2.4x}
\end{equation*}

@@

```{r, echo = FALSE}
x <- seq(0,1,by = 0.001)
bre <- function(x) x * log(x/(1-x)) + (1-x) * log((1-x)/x)
plot(x,bre(x), type = "l")
lines(x, log(1/(2.4*x)), col = "red")
abline(v = 0.15, lty = 2)
```

@@

Thus, following Kaufmann et al. (2016), define the complexity $\mathcal{K}_C(\nu)$ for the fixed confidence setting as

\begin{equation}
\mathcal{K}_C(\nu) = \inf_{A \, \text{PAC}} \lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\nu}[\tau_{\delta}]}{\log \frac{1}{\delta}}.
\end{equation}

@@

\begin{equation*}
\mathcal{K}_C(\nu) \geq \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})}
\end{equation*}

@@

\begin{equation}
\mathcal{K}_B(\nu) = \inf_{A \, \text{consistent}} \big(\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu)\big)^{-1}.
\end{equation}

Given this definition, one can interpret the bound given in Theorem 12 of Kaufmann et al. directly as a lower bound on the complexity $\mathcal{K}_B(\nu)$ (note the inverse).

@@

In contrast, Kaufmann et al. (2016) are able to show that their complexity measure derived for the fixed-confidence setting, $\mathcal{K}_C(\nu)$, can be upper-bounded using the upper bound of the sample complexity (number of pulls) given for the KL-LUCB algorithm in Kaufmann and Kalyanakrishnan (2013). Consequently, in the case of exponential family bandits, the complexity $\mathcal{K}_C(\nu)$ is bounded as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m} \frac{1}{K(\theta_a,\theta_{[m]})} \leq \mathcal{K}_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

@@

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{a=1}^K w_a \text{\em{KL}}(\mu_a, \lambda_a)\right) .
\end{equation}
\end{theorem}

And so we see that the two bounds on the sample complexity (one for Top-$m$, one for best arm identification) differ only in the way that they build the sum of Kullback-Leibler divergences over the $K$ arms. In contrast to Theorem \ref{theorem:KaufmannEtAl2016Theorem1}, here the contribution of each arm through its KL-divergence is weighted. The difference is that while we were able to very clearly describe the change of distribution in the Top-$m$ case for each arm -- for each arm, we specified an alternative model that yields a new set of $m$ optimal arms -- we now consider all alternative models that yield a different best arm than the model $\bm{\mu}$. In the previous theorem, we sum over the complexity of each individual arm, as the change in distribution only ever touches one arm, thus leading to a KL-divergence of 0 for the other arms while still changing the optimal set. Here, we consider all instances of alternative models $\bm{\lambda}$, and thus also those in which several arms are modified (and thus have a KL-divergence unequal 0 in the result of Lemma 1).

@@

\begin{equation*}
Z_{a,b}(t) = - Z_{b,a} = N_a(t) \KL(\hat{\mu},\hat{\mu}_{a,b})) + N_b(t) \KL(\hat{\mu}_{b(t)}, \hat{\mu}_{a,b}),
\end{equation*}

@@

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big)
\end{equation*}.

>>>>

\begin{equation}\label{bre_approx}
\forall x \in [0,1], \quad \der(x,1-x) = \der(1-x,x) \geq \log \frac{1}{2.4x}
\end{equation}

@@

```{r, echo = FALSE}
x <- seq(0,1,by = 0.001)
bre <- function(x) x * log(x/(1-x)) + (1-x) * log((1-x)/x)
plot(x,bre(x), type = "l", ylab = "Binary Relative Entropy")
lines(x, log(1/(2.4*x)), col = "red")
abline(v = 0.15, lty = 2)
```

@@

Thus, following Kaufmann et al. (2016), define the complexity $\kappa_C(\nu)$ for the fixed confidence setting as

\begin{equation}
\kappa_C(\nu) = \inf_{A \, \text{PAC}} \lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\nu}[\tau_{\delta}]}{\log \frac{1}{\delta}}.
\end{equation}

@@

\begin{equation*}
\kappa_C(\nu) \geq \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})}
\end{equation*}

@@

\begin{equation}
\kappa_B(\nu) = \inf_{A \, \text{consistent}} \big(\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu)\big)^{-1}.
\end{equation}

Given this definition, one can interpret the bound given in Theorem 12 of Kaufmann et al. directly as a lower bound on the complexity $\kappa_B(\nu)$ (note the inverse).

@@

In contrast, Kaufmann et al. (2016) are able to show that their complexity measure derived for the fixed-confidence setting, $\kappa_C(\nu)$, can be upper-bounded using the upper bound of the sample complexity (number of pulls) given for the KL-LUCB algorithm in Kaufmann and Kalyanakrishnan (2013). Consequently, in the case of exponential family bandits, the complexity $\kappa_C(\nu)$ is bounded as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m} \frac{1}{K(\theta_a,\theta_{[m]})} \leq \kappa_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

@@

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{a=1}^K w_a \text{\em{KL}}(\mu_a, \lambda_a)\right).
\end{equation}
\end{theorem}

And so we see that the two bounds on the sample complexity (one for Top-$m$, one for best arm identification) differ only in the way that they build the sum of Kullback-Leibler divergences over the $K$ arms. In contrast to Theorem \ref{theorem:KaufmannEtAlTheorem4}, here the contribution of each arm through its KL-divergence is weighted. The difference is that while we were able to very clearly describe the change of distribution in the Top-$m$ case for each arm -- for each arm, we specified an alternative model that yields a new set of $m$ optimal arms -- we now consider all alternative models that yield a different best arm than the model $\bm{\mu}$. In the previous theorem, we sum over the complexity of each individual arm, as the change in distribution only ever touches one arm, thus leading to a KL-divergence of 0 for the other arms while still changing the optimal set. Here, we consider all instances of alternative models $\bm{\lambda}$, and thus also those in which several arms are modified (and thus have a KL-divergence unequal 0 in the result of Lemma 1).

@@

\begin{equation*}
Z_{a,b}(t) = - Z_{b,a} = N_a(t) \KL(\hat{\mu}_a(t),\hat{\mu}_{a,b}(t)) + N_b(t) \KL(\hat{\mu}_{b}(t), \hat{\mu}_{a,b}(t)),
\end{equation*}

@@

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big).
\end{equation*}

With this, we can state the final theorem of the chapter, which gives the upper bound matching the lower bound in Theorem \ref{theorem:GarivierKaufmannTheorem1} for small risk values $\delta$ using the D-Tracking rule and Chernoff's stopping rule with the exploration term given in the previous equation:

\begin{theorem}[Theorem 14, Garivier and Kaufmann, 2016] \label{theorem:GarivierAndKaufmannTheorem14}
Let $\bm{\mu}$ be an exponential family bandit model. Let $\alpha \in [1, e/2]$ and $r(t) = O(t^{\alpha})$. Using Chernoff's stopping rule with $\beta(t,\delta) = \log(r(t)/\delta)$ and the sampling rule C-Tracking or D-Tracking,

\begin{equation*}
\lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\bm{\mu}}[\tau_{\delta}]}{\log(1/\delta)} \leq \alpha T^*(\mu).
\end{equation*}
\end{theorem}

### Conclusion of Complexity of Best Arm Identification

Overall, the two papers presented in this chapter make significant contributions to identifying the top-$1$ or the top-$m$ arms in a multi-armed bandit setting. Their contribution is primarily based on including the Kullback-Leibler divergence in the derivation of bounds and the matching algorithms and thus going beyond a sum of squared gaps which is most practical only in the case of Gaussian distributions. While the contribution in the fixed-budget setting was limited to the two-armed bandit case, the authors derive general lower bounds in the fixed-confidence setting which are primarily based on Lemma \ref{theorem:KaufmannEtAlLemma1}. Lemma \ref{theorem:KaufmannEtAlLemma1} is able to measure the "transportation cost" (see Garivier and Kaufmann, 2016) of moving from one distribution to another one in terms of Kullback-Leibler divergence. Using the information from the lower bound of how well any $\delta$-PAC algorithm can potentially be, the authors design algorithms matching these lower bounds. In the case of Top-$m$ identification, a simple application of the existing KL-LUCB algorithm works. In the case of best arm identification, the lower bound provides a way of calculating the optimal proportions in which the arms need to be pulled. Consequently, a matching algorithm can be set up so that it draws arms in these proportions, and so that it is robust against estimation errors of the plug-in estimates. Combining this sampling rule with an appropriate stopping rule makes for a $\delta$-PAC algorithm that matches the lower bound, where the upper bound in Theorem \ref{theorem:GarivierAndKaufmannTheorem14} matches the lower bound from Theorem \ref{theorem:GarivierKaufmannTheorem1} using equation \eqref{bre_approx}.

### 2017-05-28 Complexity of Best Arm Identification/kl_complexity.tex
Add conclusion to KL Complexity chapter

Thus, following Kaufmann et al. (2016), define the complexity
\(\mathcal{K}_C(\nu)\) for the fixed confidence setting as

\begin{equation}
\mathcal{K}_C(\nu) = \inf_{A \, \text{PAC}} \lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\nu}[\tau_{\delta}]}{\log \frac{1}{\delta}}.
\end{equation}

@@

\begin{equation}
\mathcal{K}_B(\nu) = \inf_{A \, \text{consistent}} \big(\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu)\big)^{-1}.
\end{equation}

Given this definition, one can interpret the bound given in Theorem 12
of Kaufmann et al. directly as a lower bound on the complexity
\(\mathcal{K}_B(\nu)\) (note the inverse).

@@

In contrast, Kaufmann et al. (2016) are able to show that their
complexity measure derived for the fixed-confidence setting,
\(\mathcal{K}_C(\nu)\), can be upper-bounded using the upper bound of
the sample complexity (number of pulls) given for the KL-LUCB algorithm
in Kaufmann and Kalyanakrishnan (2013). Consequently, in the case of
exponential family bandits, the complexity \(\mathcal{K}_C(\nu)\) is
bounded as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m} \frac{1}{K(\theta_a,\theta_{[m]})} \leq \mathcal{K}_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

@@

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{a=1}^K w_a \text{\em{KL}}(\mu_a, \lambda_a)\right) .
\end{equation}
\end{theorem}

And so we see that the two bounds on the sample complexity (one for
Top-\(m\), one for best arm identification) differ only in the way that
they build the sum of Kullback-Leibler divergences over the \(K\) arms.
In contrast to Theorem \ref{theorem:KaufmannEtAl2016Theorem1}, here the
contribution of each arm through its KL-divergence is weighted. The
difference is that while we were able to very clearly describe the
change of distribution in the Top-\(m\) case for each arm -- for each
arm, we specified an alternative model that yields a new set of \(m\)
optimal arms -- we now consider all alternative models that yield a
different best arm than the model \(\bm{\mu}\). In the previous theorem,
we sum over the complexity of each individual arm, as the change in
distribution only ever touches one arm, thus leading to a KL-divergence
of 0 for the other arms while still changing the optimal set. Here, we
consider all instances of alternative models \(\bm{\lambda}\), and thus
also those in which several arms are modified (and thus have a
KL-divergence unequal 0 in the result of Lemma 1).

@@

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big)
\end{equation*}

.

>>>>

Thus, following Kaufmann et al. (2016), define the complexity
\(\kappa_C(\nu)\) for the fixed confidence setting as

\begin{equation}
\kappa_C(\nu) = \inf_{A \, \text{PAC}} \lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\nu}[\tau_{\delta}]}{\log \frac{1}{\delta}}.
\end{equation}

@@

\begin{equation}
\kappa_B(\nu) = \inf_{A \, \text{consistent}} \big(\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu)\big)^{-1}.
\end{equation}

Given this definition, one can interpret the bound given in Theorem 12
of Kaufmann et al. directly as a lower bound on the complexity
\(\kappa_B(\nu)\) (note the inverse).

@@

In contrast, Kaufmann et al. (2016) are able to show that their
complexity measure derived for the fixed-confidence setting,
\(\kappa_C(\nu)\), can be upper-bounded using the upper bound of the
sample complexity (number of pulls) given for the KL-LUCB algorithm in
Kaufmann and Kalyanakrishnan (2013). Consequently, in the case of
exponential family bandits, the complexity \(\kappa_C(\nu)\) is bounded
as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m} \frac{1}{K(\theta_a,\theta_{[m]})} \leq \kappa_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

@@

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{a=1}^K w_a \text{\em{KL}}(\mu_a, \lambda_a)\right).
\end{equation}
\end{theorem}

And so we see that the two bounds on the sample complexity (one for
Top-\(m\), one for best arm identification) differ only in the way that
they build the sum of Kullback-Leibler divergences over the \(K\) arms.
In contrast to Theorem \ref{theorem:KaufmannEtAlTheorem4}, here the
contribution of each arm through its KL-divergence is weighted. The
difference is that while we were able to very clearly describe the
change of distribution in the Top-\(m\) case for each arm -- for each
arm, we specified an alternative model that yields a new set of \(m\)
optimal arms -- we now consider all alternative models that yield a
different best arm than the model \(\bm{\mu}\). In the previous theorem,
we sum over the complexity of each individual arm, as the change in
distribution only ever touches one arm, thus leading to a KL-divergence
of 0 for the other arms while still changing the optimal set. Here, we
consider all instances of alternative models \(\bm{\lambda}\), and thus
also those in which several arms are modified (and thus have a
KL-divergence unequal 0 in the result of Lemma 1).

@@

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big).
\end{equation*}

With this, we can state the final theorem of the chapter, which gives
the upper bound matching the lower bound in Theorem
\ref{theorem:GarivierKaufmannTheorem1} for small risk values \(\delta\)
using the D-Tracking rule and Chernoff's stopping rule with the
exploration term given in the previous equation:

\begin{theorem}[Theorem 14, Garivier and Kaufmann, 2016] \label{theorem:GarivierAndKaufmannTheorem14}
Let $\bm{\mu}$ be an exponential family bandit model. Let $\alpha \in [1, e/2]$ and $r(t) = O(t^{\alpha})$. Using Chernoff's stopping rule with $\beta(t,\delta) = \log(r(t)/\delta)$ and the sampling rule C-Tracking or D-Tracking,

\begin{equation*}
\lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\bm{\mu}}[\tau_{\delta}]}{\log(1/\delta)} \leq \alpha T^*(\mu).
\end{equation*}
\end{theorem}

\subsubsection{Conclusion of Complexity of Best Arm
Identification}\label{conclusion-of-complexity-of-best-arm-identification}

Overall, the two papers presented in this chapter make significant
contributions to identifying the top-\(1\) or the top-\(m\) arms in a
multi-armed bandit setting. Their contribution is primarily based on
including the Kullback-Leibler divergence in the derivation of bounds
and the matching algorithms and thus going beyond a sum of squared gaps
which is most practical only in the case of Gaussian distributions.
While the contribution in the fixed-budget setting was limited to the
two-armed bandit case, the authors derive general lower bounds in the
fixed-confidence setting which are primarily based on Lemma
\ref{theorem:KaufmannEtAlLemma1}. Lemma \ref{theorem:KaufmannEtAlLemma1}
is able to measure the ``transportation cost'' (see Garivier and
Kaufmann, 2016) of moving from one distribution to another one in terms
of Kullback-Leibler divergence. Using the information from the lower
bound of how well any \(\delta\)-PAC algorithm can potentially be, the
authors design algorithms matching these lower bounds. In the case of
Top-\(m\) identification, a simple application of the existing KL-LUCB
algorithm works. In the case of best arm identification, the lower bound
provides a way of calculating the optimal proportions in which the arms
need to be pulled. Consequently, a matching algorithm can be set up so
that it draws arms in these proportions, and so that it is robust
against estimation errors of the plug-in estimates. Combining this
sampling rule with an appropriate stopping rule makes for a
\(\delta\)-PAC algorithm that matches the lower bound, where the upper
bound in Theorem \ref{theorem:GarivierAndKaufmannTheorem14} matches the
lower bound from Theorem \ref{theorem:GarivierKaufmannTheorem1} using
equation \eqref{bre_approx}.
